---
title: "A small example of scABC"
author: "Timothy Daley"
date: "8/27/2017"
output: html_document
---

Here we'll walk through an example analysis of a small dataset, 25 cells of each of BJ and TF1 cell lines.  

The bam files were obtained by applying the Kundaje pipeline (https://github.com/kundajelab/atac_dnase_pipelines) to all of the single cells from the six cell lines described in the paper.  

# Detailed workflow

```{r}
# always run
library(devtools)
devtools::install_github("timydaley/scABC", force = TRUE)
library(scABC)
```
```{r eval=FALSE}
# not evaluated, preprocessing
bamfiles = paste0("SRX860", c(379:474, 187:282), "Chr12.bam")
peakfile = "../inst/extdata/chr12Peaks.bed"
peaks = selectPeaks(peakfile)
ForeGround = getCountsMatrix(bamfiles, peaks)
ForeGroundFiltered = filterPeaks(ForeGround$ForeGroundMatrix, peaks)
peaks = ForeGroundFiltered$peaks
BackGround = getBackground(bamfiles, peaks)
ForeGroundBackGroundFiltered = filterSamples(ForeGround = ForeGroundFiltered$ForeGroundMatrix, 
                                             BackGround = BackGround$BackGroundMatrix)
ForeGroundMatrix = ForeGroundBackGroundFiltered$ForeGroundMatrix
BackGroundMatrix = ForeGroundBackGroundFiltered$BackGroundMatrix
```

Now that the data is preprocessed we can apply scABC to cluster it.

```{r}
data(testData, package = "scABC")
nClusters = 1:10
BackGroundMedian = apply(testData$BackGroundMatrix, 2, median)
GapStat = getGapStat(testData$ForeGroundMatrix, BackGroundMedian, 
                     nClusters=nClusters, quiet = TRUE)
GapStat$nClusterOptimal
plotGapStat(GapStat, nClusters = nClusters, main = "TF1 vs BJ")
```

2 clusters is optimal, as it should be.  We'll compute the landmarks using 2 clusters.

```{r}
LandMarks = computeLandmarks(ForeGround = testData$ForeGroundMatrix, 
                             BackGround = testData$BackGroundMatrix, 
                             nCluster = 2, nTop = 500)
cor(LandMarks, LandMarks, method = 'spearman')
LandMarkAssignments = assign2landmarks(testData$ForeGroundMatrix, LandMarks)
```

The above steps can also be performed simply with the command
```{r eval=FALSE}
s = scABC(bamfiles, peakfile)
```

## Quality of clustering

Let's look at how good the clustering was.

```{r}
Cell2LandmarkCorrelation = cbind(apply(testData$ForeGroundMatrix, 2, function(x) cor(x, LandMarks[,1], method = 'spearman')), 
                                 apply(testData$ForeGroundMatrix, 2, function(x) cor(x, LandMarks[,2], method = 'spearman')))
library(gplots) 
library(RColorBrewer)
library(devtools)
source_url("https://raw.githubusercontent.com/obigriffith/biostar-tutorials/master/Heatmaps/heatmap.3.R")
scalered <- colorRampPalette(c("white", "red"), space = "rgb")(256)
rcols1 = brewer.pal(3, "Accent")[1:2]
rowcols1 = rcols1[testData$trueLabels[which(testData$bamfiles %in% colnames(testData$ForeGroundMatrix))]]
rcols2 = brewer.pal(3, "Dark2")[1:2]
rowcols2 = rcols2[LandMarkAssignments]
rowcols = rbind(rowcols1, rowcols2)
rownames(rowcols) = c("cell type", "cluster")
heatmap.3(Cell2LandmarkCorrelation, dendrogram='none', Rowv=FALSE, Colv=FALSE,
          trace='none', col = scalered, margin = c(5, 5), density.info = "none", 
          RowSideColors = rowcols, RowSideColorsSize=2, symm=F,symkey=F,
          symbreaks=F, scale="none")
legend("bottomleft", legend = c(unique(sapply(testData$trueLabels, toString)), paste0("cluster ", 1:2)), col = c(rcols1, rcols2), border=FALSE, bty="n", y.intersp = 0.7, cex=0.7, pch = 15)
```

## Cluster specific peaks

Let's now look at cluster specific peaks.

```{r}
PeakSelection = getClusterSpecificPvalue(ForeGround=testData$ForeGroundMatrix, cluster_assignments = LandMarkAssignments, background_medians = BackGroundMedian)
PeakPvals = PeakSelection$pvalue
head(PeakPvals)
rownames(rowcols) = c("rep info", "cluster")
x = t(testData$ForeGroundMatrix)
d = as.dist(1 - cor(x, method = "spearman"))
col.clus = hclust(d, method ="centroid")
#truncate to see patterns
x[which(x > 5)] = 5
heatmap.3(x, dendrogram='none', Rowv=FALSE, Colv=col.clus, trace='none', col = scalered, 
          margin = c(5, 5), density.info = "none", RowSideColors = rowcols, 
          RowSideColorsSize=2, main = "All peaks")
legend("bottomleft", legend = c(unique(sapply(testData$trueLabels, toString)), paste0("cluster ", 1:2)), col = c(rcols1, rcols2), border=FALSE, bty="n", y.intersp = 0.7, cex=0.7, pch = 15)
```

It's difficult to see which peaks are cluster specific because of the sparsity of the data.  Let's look at only the cluster specific peaks.

```{r}
ClusterSpecificPeaks = which(apply(PeakPvals, 1, min) < 1e-6)
length(ClusterSpecificPeaks)
 x = t(testData$ForeGroundMatrix[ClusterSpecificPeaks, ])
d = as.dist(1 - cor(x, method = "spearman"))
col.clus = hclust(d, method ="complete")
x[which(x > 5)] = 5
heatmap.3(x, dendrogram='none', Rowv=FALSE, Colv=col.clus, trace='none', col = scalered, 
          margin = c(5, 5), density.info = "none", RowSideColors = rowcols, 
          RowSideColorsSize=2, main = "Cluster specific peaks")
legend("bottomleft", legend = c(unique(sapply(testData$trueLabels, toString)), paste0("cluster ", 1:2)), col = c(rcols1, rcols2), border=FALSE, bty="n", y.intersp = 0.7, cex=0.7, pch = 15)
```

It seems that most of the cluster specific peaks are specific to BJ, which is consistent with results from analysing all 6 cell lines together.
