---
title: "Check batch effect"
author: "Zhixiang Lin"
date: "3/6/2017"
output: pdf_document
---

How bad is the batch effect? Think about GC content, etc. Need the batch information from Mahdi.

```{r}
library(readr)
library(WeightedCluster)
```

## load data, K562

data matrix

```{r}
ForeGround <- read_csv("/Users/zhixianglin/Documents/collaboration/mahdi/scATAC/K562/416Cells/ForeGround.csv", col_names = FALSE)
ForeGround <- as.matrix(ForeGround)

BackGround <- read_csv("/Users/zhixianglin/Documents/collaboration/mahdi/scATAC/K562/416Cells/BackGround.csv", col_names = FALSE)
BackGround <- as.matrix(BackGround)
```

batch information

```{r}
batch <- read_delim("/Users/zhixianglin/Documents/collaboration/mahdi/scATAC/K562/416Cells/SampleOrderBrief.txt", "\t", escape_double = FALSE, col_names = FALSE, trim_ws = TRUE)
batch <- batch[,2][[1]]
print(table(batch))
```

## run weighted K-medoid

Calculate 1 - Spearman 

```{r}
distS <- 1-cor(ForeGround, method="spearman")
```

Calculate the median of BackGround for each sample

```{r}
BackGroundMedian <- apply(BackGround, 2, median)
```

Boxplot of BackGroundMedian stratified by batch

```{r}
boxplot(BackGroundMedian~batch)
```

```{r}
nCluster <- 6
lambda <- 1
a <- quantile(BackGroundMedian, 0.5)
W <- 1/(1+exp(-lambda*(BackGroundMedian-a)))
resultW <- wcKMedoids(distS, k=nCluster, weights=W)
clusterW <- resultW$clustering
clusterW <- as.numeric(factor(clusterW))
```

## get landmark 

$landmark$ is $num of peaks \times num of lanmarks$

```{r}
landmarks <- c()
  for (i in 1:nCluster){
    tmp <- which(clusterW==i)
    if (length(tmp)==1){
        landmarks <- cbind(landmarks, ForeGround[,tmp]  )  
    } else {
        landmarks <- cbind(landmarks, rowSums(ForeGround[,tmp])  ) 
    }
  }
```


```{r}
selectTop <- function(x, top){
  thres <- sort(x, decreasing=T)[top]
  x[x<thres] <- 0
  return(x)
}
```

pick the top peaks in the landmarks

```{r}
top <- 2000
landmarksTop <- apply(landmarks, 2, selectTop, top)
```

## run KNN  

```{r}
scor <- cor(ForeGround, landmarksTop, method="spearman")
clusterWKNN <- apply(scor, 1, which.max) 
```


## check the clustering result.

```{r}
cells <- unique(batch)
```

```{r}
getClusterCount <- function(cluster, samples, cells){
  ### input
  #cluster: clustering result for each sample
  #samples: vector with the cell types for each sample
  #cells: order of the cell types
  clusterCount <- matrix(0, nrow=length(unique(cluster)), ncol=length(cells))
  for (i in 1:length(unique(cluster))){
    tmp <- samples[which(cluster==i)]
    for (j in 1:length(cells)){
      cell <- cells[j]
      clusterCount[i, j] <- sum(tmp==cell)
    }
  }
  tmp <- apply(clusterCount, 2, which.max)
  if (max(table(tmp))>1){
    seqs <- c()
    for (i in 1:nrow(clusterCount)){
      if (i==1){
        seqs <- c(seqs, which.max(clusterCount[,i])[1])
      } 
      if (i>1 & i<nrow(clusterCount)){
        seqs <- c(seqs, c(c(1:nrow(clusterCount))[-seqs])[which.max(clusterCount[-seqs,i])[1]])  
      }
      if (i==nrow(clusterCount)){
        seqs <- c(seqs, c(1:nrow(clusterCount))[-seqs]) 
      }
    }
  } else {
    seqs <- tmp
  }
  clusterCount <- clusterCount[seqs,]
  row.names(clusterCount) <- paste("cluster", 1:length(unique(cluster)))
  colnames(clusterCount) <- cells
  return(clusterCount)
}

getCorrectCount <- function(clusterCount){
  ### input
  #c lusterCount: output from getClusterCount
  ### output, we assign the cell type of each cluster by the majority
  # c(the number of correctly clusterd cells, the percentage of correctly clustered cells)
  return(c( sum(apply(clusterCount, 1, max)),  sum(apply(clusterCount, 1, max))/sum(clusterCount)) )
}
```

weighted K-medoids

```{r}
ClusterCountW <- getClusterCount(cluster=clusterW, samples=batch, cells=cells)
#CorrectCountW <- getCorrectCount(ClusterCountW)
print(ClusterCountW)
#print(CorrectCountW)
```

weighted K-medoids + KNN

```{r}
ClusterCountWKNN <- getClusterCount(cluster=clusterWKNN, samples=batch, cells=cells)
#CorrectCountWKNN <- getCorrectCount(ClusterCountWKNN)
print(ClusterCountWKNN)
```

## change number of clusters to 2

```{r}
nCluster <- 2
lambda <- 1
a <- quantile(BackGroundMedian, 0.5)
W <- 1/(1+exp(-lambda*(BackGroundMedian-a)))
resultW <- wcKMedoids(distS, k=nCluster, weights=W)
clusterW <- resultW$clustering
clusterW <- as.numeric(factor(clusterW))
landmarks <- c()
for (i in 1:nCluster){
    tmp <- which(clusterW==i)
    if (length(tmp)==1){
        landmarks <- cbind(landmarks, ForeGround[,tmp]  )  
    } else {
        landmarks <- cbind(landmarks, rowSums(ForeGround[,tmp])  ) 
    }
}
top <- 2000
landmarksTop <- apply(landmarks, 2, selectTop, top)
scor <- cor(ForeGround, landmarksTop, method="spearman")
clusterWKNN <- apply(scor, 1, which.max) 
ClusterCountW <- getClusterCount(cluster=clusterW, samples=batch, cells=cells)
print(ClusterCountW)
ClusterCountWKNN <- getClusterCount(cluster=clusterWKNN, samples=batch, cells=cells)
print(ClusterCountWKNN)
```

```{r}
nCluster <- 3
lambda <- 1
a <- quantile(BackGroundMedian, 0.5)
W <- 1/(1+exp(-lambda*(BackGroundMedian-a)))
resultW <- wcKMedoids(distS, k=nCluster, weights=W)
clusterW <- resultW$clustering
clusterW <- as.numeric(factor(clusterW))
landmarks <- c()
for (i in 1:nCluster){
    tmp <- which(clusterW==i)
    if (length(tmp)==1){
        landmarks <- cbind(landmarks, ForeGround[,tmp]  )  
    } else {
        landmarks <- cbind(landmarks, rowSums(ForeGround[,tmp])  ) 
    }
}
top <- 2000
landmarksTop <- apply(landmarks, 2, selectTop, top)
scor <- cor(ForeGround, landmarksTop, method="spearman")
clusterWKNN <- apply(scor, 1, which.max) 
ClusterCountW <- getClusterCount(cluster=clusterW, samples=batch, cells=cells)
print(ClusterCountW)
ClusterCountWKNN <- getClusterCount(cluster=clusterWKNN, samples=batch, cells=cells)
print(ClusterCountWKNN)
```