---
title: "Explore_scATACSeq_6celllines-3"
author: "Zhixiang Lin"
date: "2/28/2017"
output: pdf_document
---

This version implements the k-means algorithm

## Explore data

Load the data

```{r}
library(readr)
library(gplots)
library(Heatplus)
library(vegan)
library(RColorBrewer)
library(rafalib)
```

```{r, eval=FALSE}
data <- read_csv("~/Documents/collaboration/mahdi/scATAC/SelectedPeaks/ForeGround.csv",col_names = FALSE)
data <- as.matrix(data)
save(data, file="~/Documents/collaboration/mahdi/scATAC/SelectedPeaks/ForeGround.rda")
set.seed(123)
sub <- sample(nrow(data), 10000)
data_sub <- data[sub,]
save(data_sub, file="~/Documents/collaboration/mahdi/scATAC/SelectedPeaks/ForeGround_sub.rda")
```

```{r, eval=FALSE}
load("~/Documents/collaboration/mahdi/scATAC/SelectedPeaks/ForeGround.rda")
load("~/Documents/collaboration/mahdi/scATAC/SelectedPeaks/ForeGround_sub.rda")
samples <- read_delim("~/Documents/collaboration/mahdi/scATAC/SelectedPeaks/SampleOrder.txt", 
    "\t", escape_double = FALSE, col_names = FALSE, trim_ws = TRUE)
samples <- samples[,2][[1]]
table(samples)
```

```{r}
load("C:/Users/Zhixiang/Data/CellLines6/updated/ForeGround.rda")
load("C:/Users/Zhixiang/Data/CellLines6/updated/ForeGround_sub.rda")
samples <- read_delim("C:/Users/Zhixiang/Data/CellLines6/updated/SampleOrder.txt", 
    "\t", escape_double = FALSE, col_names = FALSE, trim_ws = TRUE)
samples <- samples[,2][[1]]
table(samples)
```

```{r}
hist(colSums(data!=0), breaks=100)
```

## spearman correlation

all samples

```{r}
distSsub <- 1-cor(data_sub, method="spearman")
distS <- 1-cor(data, method="spearman")
colnames(distS) <- samples
rownames(distS) <- samples

cells <- unique(samples)
colcols <- rep(NA, length(samples))
for (i in 1:length(cells)){
  cell <- cells[i]
  colcols[which(samples==cell)] <- brewer.pal(6, "Dark2")[i]
}
col.clus = hclust(as.dist(distS), method="complete")
```

```{r}
scaleyellowred <- colorRampPalette(c("lightyellow", "red"), space = "rgb")(100)
heatmap.2(distS, symm = TRUE, dendrogram="column", Colv = as.dendrogram(col.clus), col = scaleyellowred, ColSideColors = colcols, margins = c(10, 3),  trace = "none", density.info = "none", labRow = "", labCol = "")
```

Get the legends

```{r}
plot(1:6, 1:6, col=brewer.pal(6, "Dark2"), type="n")
legend("topleft", legend = cells, col = brewer.pal(6, "Dark2"), lwd = 10,bty = "n")
```

subset of regions

```{r}
distSsub <- 1-cor(data_sub, method="spearman")
colnames(distSsub) <- samples
rownames(distSsub) <- samples

cells <- unique(samples)
colcols <- rep(NA, length(samples))
for (i in 1:length(cells)){
  cell <- cells[i]
  colcols[which(samples==cell)] <- brewer.pal(6, "Dark2")[i]
}
col.clus = hclust(as.dist(distSsub), method="complete")
```

```{r}
scaleyellowred <- colorRampPalette(c("lightyellow", "red"), space = "rgb")(100)
heatmap.2(distSsub, symm = TRUE, dendrogram="column", Colv = as.dendrogram(col.clus), col = scaleyellowred, ColSideColors = colcols, margins = c(10, 3),  trace = "none", density.info = "none", labRow = "", labCol = "")
```

a permutation of the samples, subset of regions

```{r}
set.seed(123)
perm <- sample(ncol(data_sub))
distSsubPerm <- distSsub[perm, perm]
colcolsPerm <- colcols[perm]
col.clusPerm = hclust(as.dist(distSsubPerm), method="complete")
scaleyellowred <- colorRampPalette(c("lightyellow", "red"), space = "rgb")(100)
heatmap.2(distSsubPerm, symm = TRUE, dendrogram="column", Colv = as.dendrogram(col.clusPerm), col = scaleyellowred, ColSideColors = colcolsPerm, margins = c(10, 3),  trace = "none", density.info = "none", labRow = "", labCol = "")
```


a subset of samples, pick 60 cells per cell line

```{r}
set.seed(123)
num <- 60
subSseq <- c()
for (i in 1:length(cells)){
  cell <- cells[i]
  subSseq <- c(subSseq, sample(which(samples==cell), num) )
}

datasubS <- data[,subSseq]
samplessubS <- samples[subSseq]
distSsubS <- distS[subSseq, subSseq]
colcolssubS <- colcols[subSseq]
col.clussubS = hclust(as.dist(distSsubS), method="complete")

heatmap.2(distSsubS, symm = TRUE, dendrogram="column", Colv = as.dendrogram(col.clussubS), col = scaleyellowred, ColSideColors = colcolssubS, margins = c(10, 3),  trace = "none", density.info = "none", labRow = "", labCol = "")
```

add those problemetic samples

```{r}
set.seed(123)
num <- 60
pseq <- col.clus$order[c(136:149,932:946)]
datatmp <- data[,-pseq]
samplestmp <- samples[-pseq]
datasubS <- data[,pseq]
samplessubS <- samples[pseq]

subSseq <- c()
for (i in 1:length(cells)){
  cell <- cells[i]
  numtmp <- num - sum(samplessubS==cell)
  subSseq <- c(subSseq, sample(which(samplestmp==cell), numtmp) )
}
datasubS <- cbind(datasubS, datatmp[,subSseq])
samplessubS <- c(samplessubS, samplestmp[subSseq])

distSsubS <- 1-cor(datasubS, method="spearman")

colcolssubS <- rep(NA, length(samplessubS))
for (i in 1:length(cells)){
  cell <- cells[i]
  colcolssubS[which(samplessubS==cell)] <- brewer.pal(6, "Dark2")[i]
}

col.clussubS = hclust(as.dist(distSsubS), method="complete")

heatmap.2(distSsubS, symm = TRUE, dendrogram="column", Colv = as.dendrogram(col.clussubS), col = scaleyellowred, ColSideColors = colcolssubS, margins = c(10, 3),  trace = "none", density.info = "none", labRow = "", labCol = "")
```

## select top peak, overlap


```{r}
caloverlap <- function(x, y, top){
  xtmp <- 0*x
  ytmp <- 0*y
  xtmp[order(x, decreasing=T)[1:top]] <- 1  
  ytmp[order(y, decreasing=T)[1:top]] <- 1 
  sum((xtmp+ytmp)==2)/top
}

getdataBin <- function(x, top){
  xtmp <- 0*x
  xtmp[order(x, decreasing=T)[1:top]] <- 1  
  return(xtmp)
}

calBinDist <- function(column, data, top){
  return(colSums(data*column)/top)
}
```

```{r}
top <- 1000
dataBin <- apply(data, 2, getdataBin, top)
distTop <- 1-crossprod(dataBin)/top
col.clus = hclust(as.dist(distTop), method="complete")
```

```{r}
scaleyellowred <- colorRampPalette(c("lightyellow", "red"), space = "rgb")(100)
heatmap.2(distTop, symm = TRUE, dendrogram="column", Colv = as.dendrogram(col.clus), col = scaleyellowred, ColSideColors = colcols, margins = c(10, 3),  trace = "none", density.info = "none", labRow = "", labCol = "")
```

```{r}
datasubSBin <- apply(datasubS, 2, getdataBin, top)
distTop <- 1-crossprod(datasubSBin)/top
col.clus = hclust(as.dist(distTop), method="complete")
scaleyellowred <- colorRampPalette(c("lightyellow", "red"), space = "rgb")(100)
heatmap.2(distTop, symm = TRUE, dendrogram="column", Colv = as.dendrogram(col.clus), col = scaleyellowred, ColSideColors = colcolssubS, margins = c(10, 3),  trace = "none", density.info = "none", labRow = "", labCol = "")
```

## k-means on the subset of data

k-means on orginal data

```{r}
datasubS_pca <- prcomp(t(datasubS))$x[,1:30]
nComp <- 5
ncluster <- 6
kmeansCluster <- kmeans(datasubS_pca[,1:nComp], centers=ncluster, nstart=5000)$cluster
for (i in 1:ncluster){
  print(paste("cluster", i))
  print(table(samplessubS[which(kmeansCluster==i)]))
}
```

k-means on orginal data, log2 scale

```{r}
datasubS_pca <- prcomp(t(log2(datasubS+1)))$x[,1:30]
nComp <- 20
ncluster <- 6
kmeansCluster <- kmeans(datasubS_pca[,1:nComp], centers=ncluster, nstart=5000)$cluster
for (i in 1:ncluster){
  print(paste("cluster", i))
  print(table(samplessubS[which(kmeansCluster==i)]))
}
```

k-means on binarized data

```{r}
datasubS_pca <- prcomp(t(datasubSBin))$x[,1:30]
nComp <- 5
ncluster <- 6
kmeansCluster <- kmeans(datasubS_pca[,1:nComp], centers=ncluster, nstart=5000)$cluster
for (i in 1:ncluster){
  print(paste("cluster", i))
  print(table(samplessubS[which(kmeansCluster==i)]))
}
```

```{r}
nComp <- 30
ncluster <- 6
kmeansCluster <- kmeans(datasubS_pca[,1:nComp], centers=ncluster, nstart=5000)$cluster
for (i in 1:ncluster){
  print(paste("cluster", i))
  print(table(samplessubS[which(kmeansCluster==i)]))
}
```

## subset of regions for faster computation

subset of regions

```{r}
distSsub <- 1-cor(data_sub, method="spearman")
colnames(distSsub) <- samples
rownames(distSsub) <- samples

cells <- unique(samples)
colcols <- rep(NA, length(samples))
for (i in 1:length(cells)){
  cell <- cells[i]
  colcols[which(samples==cell)] <- brewer.pal(6, "Dark2")[i]
}
col.clus = hclust(as.dist(distSsub), method="complete")
```

```{r}
scaleyellowred <- colorRampPalette(c("lightyellow", "red"), space = "rgb")(100)
heatmap.2(distSsub, symm = TRUE, dendrogram="column", Colv = as.dendrogram(col.clus), col = scaleyellowred, ColSideColors = colcols, margins = c(10, 3),  trace = "none", density.info = "none", labRow = "", labCol = "")
```

```{r}
distSsub <- 1-cor(data_sub, method="spearman")
colnames(distSsub) <- samples
rownames(distSsub) <- samples

cells <- unique(samples)
colcols <- rep(NA, length(samples))
for (i in 1:length(cells)){
  cell <- cells[i]
  colcols[which(samples==cell)] <- brewer.pal(6, "Dark2")[i]
}
col.clus = hclust(as.dist(distSsub), method="average")
```

```{r}
scaleyellowred <- colorRampPalette(c("lightyellow", "red"), space = "rgb")(100)
heatmap.2(distSsub, symm = TRUE, dendrogram="column", Colv = as.dendrogram(col.clus), col = scaleyellowred, ColSideColors = colcols, margins = c(10, 3),  trace = "none", density.info = "none", labRow = "", labCol = "")
```

a permutation of the samples, subset of regions

```{r}
set.seed(123)
perm <- sample(ncol(data_sub))
distSsubPerm <- distSsub[perm, perm]
colcolsPerm <- colcols[perm]
col.clusPerm = hclust(as.dist(distSsubPerm), method="complete")
scaleyellowred <- colorRampPalette(c("lightyellow", "red"), space = "rgb")(100)
heatmap.2(distSsubPerm, symm = TRUE, dendrogram="column", Colv = as.dendrogram(col.clusPerm), col = scaleyellowred, ColSideColors = colcolsPerm, margins = c(10, 3),  trace = "none", density.info = "none", labRow = "", labCol = "")
```

a subset of the samples

```{r}
set.seed(123)
num <- 50
subSseq <- c()
for (i in 1:length(cells)){
  cell <- cells[i]
  subSseq <- c(subSseq, sample(which(samples==cell), num) )
}

datasubS <- data[,subSseq]
samplessubS <- samples[subSseq]
distSsubS <- 1-cor(datasubS, method="spearman")

cells <- unique(samplessubS)
colcolssubS <- rep(NA, length(samplessubS))
for (i in 1:length(cells)){
  cell <- cells[i]
  colcolssubS[which(samplessubS==cell)] <- brewer.pal(6, "Dark2")[i]
}

col.clussubS = hclust(as.dist(distSsubS), method="average")

heatmap.2(distSsubS, symm = TRUE, dendrogram="column", Colv = as.dendrogram(col.clussubS), col = scaleyellowred, ColSideColors = colcolssubS, margins = c(10, 3),  trace = "none", density.info = "none", labRow = "", labCol = "")
```

cluster by the average option

```{r}
distS <- 1-cor(data, method="spearman")
colnames(distS) <- samples
rownames(distS) <- samples

cells <- unique(samples)
colcols <- rep(NA, length(samples))
for (i in 1:length(cells)){
  cell <- cells[i]
  colcols[which(samples==cell)] <- brewer.pal(6, "Dark2")[i]
}
col.clus = hclust(as.dist(distS), method="average")
```

```{r}
plot(col.clus)
```

```{r}
scaleyellowred <- colorRampPalette(c("lightyellow", "red"), space = "rgb")(100)
heatmap.2(distS, symm = TRUE, dendrogram="column", Colv = as.dendrogram(col.clus), col = scaleyellowred, ColSideColors = colcols, margins = c(10, 3),  trace = "none", density.info = "none", labRow = "", labCol = "")
```

## calculate the sum for each cell type

```{r}
cells <- unique(samples)
signi <- c()
for (cell in cells){
  signi <- cbind(signi, rowSums(data[,which(samples==cell)])  )
}
```

Assign cluster label

```{r}
datasigni <- cor(data, signi, method="spearman")
cluster_lab <- apply(datasigni, 1, which.max)
```

```{r}
for (i in 1:length(unique(cluster_lab))){
  print(i)
  print(table(samples[which(cluster_lab==i)]))  
}
```

Perfect cluster!!

## Use this strategy on the subset of samples

```{r}
cells1 <- unique(samplessubS)
signi1 <- c()
for (cell in cells1){
  signi1 <- cbind(signi1, rowSums(datasubS[,which(samplessubS==cell)])  )
}
```

Assign cluster label

```{r}
datasigni1 <- cor(datasubS, signi1, method="spearman")
cluster_lab1 <- apply(datasigni1, 1, which.max)
```

Assign cluster label, full dataset

```{r}
datasigni2 <- cor(data, signi1, method="spearman")
cluster_lab2 <- apply(datasigni2, 1, which.max)
for (i in 1:length(unique(cluster_lab2))){
  print(i)
  print(table(samples[which(cluster_lab2==i)]))  
}
```

```{r}
for (i in 1:length(unique(cluster_lab1))){
  print(i)
  print(table(samplessubS[which(cluster_lab1==i)]))  
}
```

## modified k-means

Does not seem to work

```{r}
kmeansSpearman <- function(data, samples, ncenters, inicenter, maxiter){
  clustercenter <- inicenter
  for (iter in 1:maxiter){
    scor <- cor(data, clustercenter, method="spearman")
    # assign cluster
    cluster <- apply(scor, 1, which.max) 
    print(iter)
    print(table(cluster))
    # recalculate clustercenter
    clustercenter <- c()
    for (i in 1:ncenters){
      tmp <- which(cluster==i)
      if (length(tmp)==1){
        clustercenter <- cbind(clustercenter, data[,tmp]  )  
      } else {
        clustercenter <- cbind(clustercenter, rowSums(data[,tmp])  ) 
      }
    }
  }
  return(cluster)
}
```

Good start

```{r}
kmeansSpearman(data=datasubS, samples=samplessubS, ncenters=6, inicenter=signi1, maxiter=50)
```

```{r}
kmeansSpearman(data=data, samples=samples, ncenters=6, inicenter=signi1, maxiter=50)
```

debug, nope, does not work

```{r}
clustercenter <- signi1
scor <- cor(data, clustercenter, method="spearman")
cluster <- apply(scor, 1, which.max) 
clustercenter.new <- c()
for (i in 1:6){
      tmp <- which(cluster==i)
      clustercenter.new <- cbind(clustercenter.new, rowSums(data[,tmp])  ) 
}
scor.new <- cor(data, clustercenter.new, method="spearman")
cluster.new <- apply(scor.new, 1, which.max) 
```

Random start. This get stucks

```{r}
set.seed(12)
signi1Random <- datasubS[,sample(ncol(datasubS), 6)]
kmeansSpearman(data=datasubS, samples=samplessubS, ncenters=6, inicenter=signi1Random, maxiter=50)
```

Another random start

```{r}
set.seed(12)
cells1 <- unique(samplessubS)
randomcluster <- sample(samplessubS)
signiR <- c()
for (cell in cells1){
  signiR <- cbind(signiR, rowSums(datasubS[,which(randomcluster==cell)])  )
}

kmeansSpearman(data=datasubS, samples=samplessubS, ncenters=6, inicenter=signiR, maxiter=50)
```

## try PAM

Spearman 

```{r}
library(cluster)
distS <- 1-cor(data, method="spearman")
resultpamS <- pam(x=distS, k=6, diss = TRUE)
clusterS <- resultpamS$clustering
```

kendall, this is slow!

```{r, eval=FALSE}
distK <- 1-cor(data, method="kendall")
resultpamK <- pam(x=distK, k=6, diss = TRUE)
clusterK <- resultpam$clustering
```


Agglomerative step afterwards

```{r}
ncenters <- length(unique(clusterS))
clustercenter <- c()
  for (i in 1:ncenters){
    tmp <- which(clusterS==i)
    if (length(tmp)==1){
        clustercenter <- cbind(clustercenter, data[,tmp]  )  
    } else {
        clustercenter <- cbind(clustercenter, rowSums(data[,tmp])  ) 
    }
}
scor <- cor(data, clustercenter, method="spearman")
clusterSA <- apply(scor, 1, which.max) 
```

check clusters

```{r}
for (i in 1:length(unique(clusterS))){
  print(i)
  print(table(samples[which(clusterS==i)]))  
}
```

```{r}
for (i in 1:length(unique(clusterSA))){
  print(i)
  print(table(samples[which(clusterSA==i)]))  
}
```

```{r}
cells <- unique(samples)
num <- 10
subSseq <- c()
for (i in 1:length(cells)){
  cell <- cells[i]
  subSseq <- c(subSseq, sample(which(samples==cell), num) )
}

datasubS <- data[,subSseq]
samplessubS <- samples[subSseq]
distSsubS <- 1-cor(datasubS, method="spearman")

resultpamSsubS <- pam(x=distSsubS, k=6, diss = TRUE)
clusterSsubS <- resultpamSsubS$clustering

ncenters <- length(unique(clusterSsubS))
clustercenter <- c()
  for (i in 1:ncenters){
    tmp <- which(clusterSsubS==i)
    if (length(tmp)==1){
        clustercenter <- cbind(clustercenter, datasubS[,tmp]  )  
    } else {
        clustercenter <- cbind(clustercenter, rowSums(datasubS[,tmp])  ) 
    }
}
scorsubS <- cor(datasubS, clustercenter, method="spearman")
clusterSsubSA <- apply(scorsubS, 1, which.max) 
print(table(clusterSsubS))
print(table(clusterSsubSA))
```
