---
title: "plot4paper"
author: "Zhixiang Lin"
date: "3/6/2017"
output: pdf_document
---

this script generate figures for publication

```{r}
library(readr)
library(gplots)
library(Heatplus)
library(vegan)
library(RColorBrewer)
library(rafalib)
library(ggdendro)
library(ggplot2)
library(gridExtra)
library(cluster)
library(WeightedCluster)
```

## the sigmoid function plot

the sigmoid function. $\frac{1}{1+\exp^{-\lambda(x-a)}}$ 

```{r}
col <- brewer.pal(6, "Set1")[2]
lambda <- 1
a <- 8
x <- seq(0, 15, 0.1)
plot(x, 1/(1+exp(-lambda*(x-a))), type="l", ylab="Weight", xlab="Sequencing depth", lwd=2.5, col=col, xaxt="n")
```

```{r, eval=FALSE}
pdf("~/Dropbox/CEGS/figures_Zhixiang/sigmoid_new.pdf", width=5, height=5)
plot(x, 1/(1+exp(-lambda*(x-a))), type="l", ylab="Weight", xlab="Sequencing depth", lwd=2.5, col=col, xaxt="n")
dev.off()
```

## heatmap for a subset of cells and top peaks(from landmark) for demonstration in figure 1

Load the data matrix, 946 cells

desktop 

```{r}
data <- read_csv("C:/Users/Zhixiang/Data/CellLines6/Landmark-Index/946Cells/ForeGround.csv", col_names = FALSE)
data <- as.matrix(data)
background <- read_csv("C:/Users/Zhixiang/Data/CellLines6/Landmark-Index/946Cells/BackGround.csv", col_names = FALSE)
background <- as.matrix(background)
samples <- read_delim("C:/Users/Zhixiang/Data/CellLines6/Landmark-Index/946Cells/SampleOrder.txt", "\t", escape_double = FALSE, col_names = FALSE, trim_ws = TRUE)
samples <- samples[,2][[1]]
```

Load sequence for the landmark 

```{r}
Landmarks <- read_csv("C:/Users/Zhixiang/Data/CellLines6/Landmark-Index/946Cells/Landmarks.csv")
```

The cells and rearrange the landmark by cells

```{r}
cells <- c("K562", "GM12878", "HL-60", "BJ", "TF-1", "H1")
tmp <- colnames(Landmarks)
seq <- c()
for (cell in cells){
  seq <- c(seq, which(tmp==cell))
}
Landmarks <- Landmarks[,seq]
Landmarks <- as.matrix(Landmarks)
```

Check whether the top peaks are ordered by read count. Yes they are

```{r}
order(rowSums(data[,which(samples==cells[1])]), decreasing=T)[1:10]
order(rowSums(data[,which(samples==cells[3])]), decreasing=T)[1:10]
```

Pick the peaks unique to each cell type

```{r}
unique_peak_seq <- list()
for (i in 1:length(cells)){
  #if (i==1){
    ## for K562
  #}
  ltmp <- Landmarks[,i]
  lother <- as.vector(Landmarks[,-i])
  unique_peak_seq[[i]] <- ltmp[which(sapply(ltmp, function(x, lother){sum(lother==x)}, lother)==0)]
}
for (i in unique_peak_seq){
  print(length(i))
}
```

Pick the top 30 peaks unique to each cell type

```{r}
top_num <- 30
top_peak_seq <- c()
for (i in 1:length(cells)){
  top_peak_seq <- c(top_peak_seq, (unique_peak_seq[[i]])[1:top_num]) 
}
```

Randomly pick 30 cells per cell type, top 30 unique peaks. datasub is peak by sample

```{r}
set.seed(123)
```

```{r}
num_cell <- 30
dataSub <- c()
for (cell in cells){
  dataSub <- cbind(dataSub, data[top_peak_seq, sample(which(samples==cell), num_cell)])
}
dataSubThres <- dataSub
cutoff <- 
dataSubThres[which(dataSubThres>=6)] <- 6
```

plot the heatmap

```{r}
cols <- rep(brewer.pal(6, "Dark2"), each=num_cell)
scaleyellowblack <- colorRampPalette(c("lightyellow", "black"), space = "rgb")(100)
heatmap.2(dataSubThres, dendrogram="none", col = scaleyellowblack, na.color="yellow",
          labRow = "", labCol = "",
          ## row and column order
          Rowv=FALSE, Colv = FALSE,  
          ## row and column side color
          #ColSideColors = colcolssubS
          RowSideColors = cols, 
          margins = c(10, 3),  trace = "none",   
          ## color key
          key=F, key.xlab="1 - Spearman", key.title="", keysize=1.2, 
          density.info = "none")
```

Pick the samples with deepest depth

```{r}
## median of background
backgroundMedian <- apply(background, 2, median)
num_cell <- 30
dataSub1 <- c()
for (cell in cells){
  seqtmp <- order(backgroundMedian[which(samples==cell)], decreasing=T)[1:num_cell]
  dataSub1 <- cbind(dataSub1, data[top_peak_seq, (which(samples==cell))[seqtmp]])
}
dataSubThres1 <- dataSub1
dataSubThres1[which(dataSubThres1>=6)] <- 6
```

plot the heatmap

```{r}
heatmap.2(dataSubThres1, dendrogram="none", col = scaleyellowblack, na.color="yellow",
          labRow = "", labCol = "",
          ## row and column order
          Rowv=FALSE, Colv = FALSE,  
          ## row and column side color
          #ColSideColors = colcolssubS
          RowSideColors = cols, 
          margins = c(10, 3),  trace = "none",   
          ## color key
          key=F, key.xlab="1 - Spearman", key.title="", keysize=1.2, 
          density.info = "none")
```


save the two plots

```{r, eval=FALSE}
pdf("C:/Users/Zhixiang/Dropbox/CEGS/figures_Zhixiang/small_heatmap_randomSample.pdf", width=5, height=5)
heatmap.2(dataSubThres, dendrogram="none", col = scaleyellowblack, na.color="yellow",
          labRow = "", labCol = "",
          ## row and column order
          Rowv=FALSE, Colv = FALSE,  
          ## row and column side color
          #ColSideColors = colcolssubS
          RowSideColors = cols, 
          margins = c(10, 3),  trace = "none",   
          ## color key
          key=F, key.xlab="1 - Spearman", key.title="", keysize=1.2, 
          density.info = "none")
dev.off()
```

```{r, eval=F}
pdf("C:/Users/Zhixiang/Dropbox/CEGS/figures_Zhixiang/small_heatmap_deepSample.pdf", width=5, height=5)
heatmap.2(dataSubThres1, dendrogram="none", col = scaleyellowblack, na.color="yellow",
          labRow = "", labCol = "",
          ## row and column order
          Rowv=FALSE, Colv = FALSE,  
          ## row and column side color
          #ColSideColors = colcolssubS
          RowSideColors = cols, 
          margins = c(10, 3),  trace = "none",   
          ## color key
          key=F, key.xlab="1 - Spearman", key.title="", keysize=1.2, 
          density.info = "none")
dev.off()
```

## landmap and all top peaks

```{r}
library(readr)
data <- read_csv("C:/Users/Zhixiang/Dropbox/tmp/Landmark/Landmark_reorder.csv", col_names = FALSE)
data <- as.matrix(data)
sampleinfor <- read_delim("C:/Users/Zhixiang/Dropbox/tmp/Landmark/KnnWithIntial.txt", "\t", escape_double = FALSE, col_names = FALSE, trim_ws = TRUE)
clusters <- sampleinfor[,2][[1]]
samples <- sampleinfor[,3][[1]]
```

```{r}
cells <- c("K562", "GM12878", "HL-60", "BJ", "TF-1", "H1")
cols <- rep(NA, length(samples))
for (i in 1:length(cells)){
  cell <- cells[i]
  cols[which(samples==cell)] <- brewer.pal(6, "Dark2")[i]
}
scaleyellowred <- colorRampPalette(c("lightyellow", "red"), space = "rgb")(100)

dataThres <- data
dataThres[which(dataThres>=6)] <- 6
```

```{r}
pdf("test.pdf", width=4.5/2.54, height=3.8/2.54)
#par(mar=rep(1,4))
heatmap.2(dataThres[,sort(sample(5058,200))], dendrogram="none", col = scaleyellowred, na.color="yellow",
          labRow = "", labCol = "",
          ## row and column order
          Rowv=FALSE, Colv = FALSE,  
          ## row and column side color
          #ColSideColors = colcolssubS
          #RowSideColors = cols, 
          margins = c(10, 3),  trace = "none",   
          ## color key
          key=F, key.xlab="1 - Spearman", key.title="", keysize=1.2, 
          density.info = "none")
dev.off()
```


```{r}
heatmap.2(dataThres, dendrogram="none", col = scaleyellowred, na.color="yellow",
          labRow = "", labCol = "",
          ## row and column order
          Rowv=FALSE, Colv = FALSE,  
          ## row and column side color
          #ColSideColors = colcolssubS
          RowSideColors = cols, 
          margins = c(10, 3),  trace = "none",   
          ## color key
          key=F, key.xlab="1 - Spearman", key.title="", keysize=1.2, 
          density.info = "none")
```

```{r}
col1 = hclust(as.dist(cor(dataThres[,tt], method="spearman")), method="average")
heatmap.2(dataThres[,tt], dendrogram="column", col = scaleyellowred, na.color="yellow",
          labRow = "", labCol = "",
          ## row and column order
          Rowv=FALSE, Colv = col1,  
          ## row and column side color
          #ColSideColors = colcolssubS
          RowSideColors = cols, 
          margins = c(10, 3),  trace = "none",   
          ## color key
          key=F, key.xlab="1 - Spearman", key.title="", keysize=1.2, 
          density.info = "none")
```

save figure

```{r, eval=FALSE}
pdf("/Users/Zhixiang/Dropbox/CEGS/figures_Zhixiang/heatmap_toppeaks.pdf", width=10, height=10)
heatmap.2(dataThres, dendrogram="none", col = scaleyellowred, na.color="yellow",
          labRow = "", labCol = "",
          ## row and column order
          Rowv=FALSE, Colv = FALSE,  
          ## row and column side color
          #ColSideColors = colcolssubS
          RowSideColors = cols, 
          margins = c(10, 3),  trace = "none",   
          ## color key
          key=F, key.xlab="1 - Spearman", key.title="", keysize=1.2, 
          density.info = "none")
dev.off()
```

save figure, black

```{r, eval=FALSE}
scaleyellowblack <- colorRampPalette(c("lightyellow", "black"), space = "rgb")(100)
pdf("/Users/Zhixiang/Dropbox/CEGS/figures_Zhixiang/heatmap_toppeaksBlack.pdf", width=10, height=10)
heatmap.2(dataThres, dendrogram="none", col = scaleyellowblack, na.color="yellow",
          labRow = "", labCol = "",
          ## row and column order
          Rowv=FALSE, Colv = FALSE,  
          ## row and column side color
          #ColSideColors = colcolssubS
          RowSideColors = cols, 
          margins = c(10, 3),  trace = "none",   
          ## color key
          key=F, key.xlab="1 - Spearman", key.title="", keysize=1.2, 
          density.info = "none")
dev.off()
```