---
title: "plots4labmeeting1"
author: "Zhixiang Lin"
date: "March 2, 2017"
output: pdf_document
---


```{r}
library(readr)
library(gplots)
library(Heatplus)
library(vegan)
library(RColorBrewer)
library(rafalib)
library(ggdendro)
library(ggplot2)
library(gridExtra)
library(cluster)
```

Load the data

```{r}
load("C:/Users/Zhixiang/Data/CellLines6/updated/ForeGround.rda")
load("C:/Users/Zhixiang/Data/CellLines6/updated/ForeGround_sub.rda")
samples <- read_delim("C:/Users/Zhixiang/Data/CellLines6/updated/SampleOrder.txt", 
    "\t", escape_double = FALSE, col_names = FALSE, trim_ws = TRUE)
samples <- samples[,2][[1]]
table(samples)
```

## spearman correlation, hclust

all samples

```{r}
distS <- 1-cor(data, method="spearman")
diag(distS) <- NaN
colnames(distS) <- samples
rownames(distS) <- samples

cells <- c("K562", "GM12878", "HL-60", "BJ", "TF-1", "H1")

colcols <- rep(NA, length(samples))
for (i in 1:length(cells)){
  cell <- cells[i]
  colcols[which(samples==cell)] <- brewer.pal(6, "Dark2")[i]
}

col.clus = hclust(as.dist(distS), method="average")
scaleyellowred <- colorRampPalette(c("red", "lightyellow"), space = "rgb")(100)
```

save figure

```{r, eval=FALSE}
pdf("/Users/Zhixiang/Dropbox/CEGS/figures_Zhixiang/heatmap_hclustALL.pdf", width=10, height=10)
heatmap.2(distS, dendrogram="column", col = scaleyellowred, na.color="yellow",
          labRow = "", labCol = "",
          ## row and column order
          Colv = as.dendrogram(col.clus), Rowv=rev(as.dendrogram(col.clus)),  
          ## row and column side color
          ColSideColors = colcols, RowSideColors = colcols, 
          margins = c(10, 3),  trace = "none",   
          ## color key
          key=T, key.xlab="1 - Spearman", key.title="", keysize=1.2, 
          density.info = "none")
dev.off()
```

a subset

```{r}
set.seed(123)
num <- 60
pseq <- col.clus$order[c(136:149,932:946)]
datatmp <- data[,-pseq]
samplestmp <- samples[-pseq]
datasubS <- data[,pseq]
samplessubS <- samples[pseq]

subSseq <- c()
for (i in 1:length(cells)){
  cell <- cells[i]
  numtmp <- num - sum(samplessubS==cell)
  subSseq <- c(subSseq, sample(which(samplestmp==cell), numtmp) )
}

datasubS <- cbind(datasubS, datatmp[,subSseq])
samplessubS <- c(samplessubS, samplestmp[subSseq])

distSsubS <- 1-cor(datasubS, method="spearman")
colcolssubS <- rep(NA, length(samplessubS))
for (i in 1:length(cells)){
  cell <- cells[i]
  colcolssubS[which(samplessubS==cell)] <- brewer.pal(6, "Dark2")[i]
}

col.clussubS = hclust(as.dist(distSsubS), method="average")
diag(distSsubS) <- NaN
colnames(distSsubS) <- samplessubS
rownames(distSsubS) <- samplessubS
scaleyellowred <- colorRampPalette(c("red", "lightyellow"), space = "rgb")(100)
```

save figure

```{r, eval=FALSE}
pdf("/Users/Zhixiang/Dropbox/CEGS/figures_Zhixiang/heatmap_hclust1.pdf", width=10, height=10)
heatmap.2(distSsubS, dendrogram="column", col = scaleyellowred, na.color="yellow",
          labRow = "", labCol = "",
          ## row and column order
          Colv = as.dendrogram(col.clussubS), Rowv=rev(as.dendrogram(col.clussubS)),  
          ## row and column side color
          ColSideColors = colcolssubS, RowSideColors = colcolssubS, 
          margins = c(10, 3),  trace = "none",   
          ## color key
          key=T, key.xlab="1 - Spearman", key.title="", keysize=1.2, 
          density.info = "none")
dev.off()
```
          
Get the legends

```{r, eval=FALSE}
pdf("/Users/Zhixiang/Dropbox/CEGS/figures_Zhixiang/legend_cells.pdf", width=10, height=10)
plot(1:6, 1:6, col=brewer.pal(6, "Dark2"), type="n", yaxt="n", xaxt="n", bty = 'n', xlab="", ylab="")
legend("topleft", legend = cells, col = brewer.pal(6, "Dark2"), lwd = 10,bty = "n")
dev.off()
```


only dendrogram, no heatmap. Failed

```{r, eval=FALSE}
#http://stackoverflow.com/questions/19926697/ggplot2-and-ggdendro-plotting-color-bars-under-the-node-leaves
df2<-data.frame(color=colcolssubS[col.clussubS$order],samples=samplessubS[col.clussubS$order])


p1<-ggdendrogram(col.clussubS, rotate=FALSE)


p2<-ggplot(df2, aes(samples,y=1,color=color))+geom_tile()+
  scale_y_continuous(expand=c(0,0))+
  theme(axis.title=element_blank(),
        axis.ticks=element_blank(),
        axis.text=element_blank(),
        legend.position="none")



gp1<-ggplotGrob(p1)
gp2<-ggplotGrob(p2)  

maxWidth = grid::unit.pmax(gp1$widths[2:5], gp2$widths[2:5])
gp1$widths[2:5] <- as.list(maxWidth)
gp2$widths[2:5] <- as.list(maxWidth)

grid.arrange(gp1, gp2, ncol=1,heights=c(4/5,1/5))
```

## use PAM, 1-spearman, all samples

```{r}
diag(distS) <- 0
resultpamS <- pam(x=distS, k=6, diss = TRUE)
clusterS <- resultpamS$clustering
```

check clusters. 5 misclassified

```{r}
clustertmp <- clusterS
clusterMembertmp <- matrix(0, nrow=6, ncol=6)
for (i in 1:length(unique(clustertmp))){
  tmp <- samples[which(clustertmp==i)]
  for (j in 1:length(cells)){
    cell <- cells[j]
    clusterMembertmp[i, j] <- sum(tmp==cell)
  }
}
## rearrange it
clusterMembertmp <- clusterMembertmp[apply(clusterMembertmp, 2, which.max),]
row.names(clusterMembertmp) <- paste("cluster", 1:6)
colnames(clusterMembertmp) <- cells
clusterMemberS <- clusterMembertmp
print(clusterMemberS)
```

Agglomerative step afterwards

```{r}
ncenters <- length(unique(clusterS))
clustercenter <- c()
  for (i in 1:ncenters){
    tmp <- which(clusterS==i)
    if (length(tmp)==1){
        clustercenter <- cbind(clustercenter, data[,tmp]  )  
    } else {
        clustercenter <- cbind(clustercenter, rowSums(data[,tmp])  ) 
    }
}
scor <- cor(data, clustercenter, method="spearman")
clusterSA <- apply(scor, 1, which.max) 
```

check clusters. 3 misclassified

```{r}
clustertmp <- clusterSA
clusterMembertmp <- matrix(0, nrow=6, ncol=6)
for (i in 1:length(unique(clustertmp))){
  tmp <- samples[which(clustertmp==i)]
  for (j in 1:length(cells)){
    cell <- cells[j]
    clusterMembertmp[i, j] <- sum(tmp==cell)
  }
}
## rearrange it
clusterMembertmp <- clusterMembertmp[apply(clusterMembertmp, 2, which.max),]
row.names(clusterMembertmp) <- paste("cluster", 1:6)
colnames(clusterMembertmp) <- cells
clusterMemberS <- clusterMembertmp
print(clusterMemberS)
```

order the samples

```{r}
orderSeq <- c()
for (cell in cells){
 orderSeq <- c(orderSeq, sample(which(samples==cell)))  
}
```

Heatmap for landmark

```{r}
scorm1 <- 1 - scor
colnames(scorm1) <- 1:6
rownames(scorm1) <- samples[orderSeq]
scaleyellowred <- colorRampPalette(c("red", "lightyellow"), space = "rgb")(100)
heatmap.2(scorm1[orderSeq,], dendrogram="none", 
          Rowv=FALSE, Colv=FALSE,
          col = scaleyellowred,
          labRow = "",
          RowSideColors = colcols[orderSeq], 
          margins = c(10, 3),  trace = "none",   
          key=T, key.xlab="1 - Spearman", key.title ="",
          density.info = "none")
```

save the plot

```{r, eval=FALSE}
pdf("/Users/Zhixiang/Dropbox/CEGS/figures_Zhixiang/heatmap_landmarkALL.pdf", width=10, height=10)
heatmap.2(scorm1[orderSeq,], dendrogram="none", 
          Rowv=FALSE, Colv=FALSE,
          col = scaleyellowred,
          labRow = "",
          RowSideColors = colcols[orderSeq], 
          margins = c(10, 3),  trace = "none",   
          key=T, key.xlab="1 - Spearman", key.title ="",
          density.info = "none")
dev.off()
```

## use PAM, 1-spearman, subset of samples

```{r}
resultpamS <- pam(x=distSsubS, k=6, diss = TRUE)
clusterS <- resultpamS$clustering
```

Agglomerative step afterwards

```{r}
ncenters <- length(unique(clusterS))
clustercenter <- c()
  for (i in 1:ncenters){
    tmp <- which(clusterS==i)
    if (length(tmp)==1){
        clustercenter <- cbind(clustercenter, datasubS[,tmp]  )  
    } else {
        clustercenter <- cbind(clustercenter, rowSums(datasubS[,tmp])  ) 
    }
}
scor <- cor(datasubS, clustercenter, method="spearman")
clusterSA <- apply(scor, 1, which.max) 
```

check clusters

```{r}
cells <- c("K562", "GM12878", "HL-60", "BJ", "TF-1", "H1")
```

```{r}
clustertmp <- clusterS
clusterMembertmp <- matrix(0, nrow=6, ncol=6)
for (i in 1:length(unique(clustertmp))){
  tmp <- samplessubS[which(clustertmp==i)]
  for (j in 1:length(cells)){
    cell <- cells[j]
    clusterMembertmp[i, j] <- sum(tmp==cell)
  }
}
## rearrange it
clusterMembertmp <- clusterMembertmp[apply(clusterMembertmp, 2, which.max),]
row.names(clusterMembertmp) <- paste("cluster", 1:6)
colnames(clusterMembertmp) <- cells
clusterMemberS <- clusterMembertmp
```

```{r}
clustertmp <- clusterSA
clusterMembertmp <- matrix(0, nrow=6, ncol=6)
for (i in 1:length(unique(clustertmp))){
  tmp <- samplessubS[which(clustertmp==i)]
  for (j in 1:length(cells)){
    cell <- cells[j]
    clusterMembertmp[i, j] <- sum(tmp==cell)
  }
}
## rearrange it
clusterMembertmp <- clusterMembertmp[apply(clusterMembertmp, 2, which.max),]
row.names(clusterMembertmp) <- paste("cluster", 1:6)
colnames(clusterMembertmp) <- cells
clusterMemberSA <- clusterMembertmp
```

No big difference

Get landmark and plot heatmap

Order the samples

```{r}
orderSeq <- c()
for (cell in cells){
 orderSeq <- c(orderSeq, sample(which(samplessubS==cell)))  
}
```

Heatmap for landmark

```{r}
scorm1 <- 1 - scor
scorm1 <- scorm1[,c(3,1,4,2,5,6)]
colnames(scorm1) <- 1:6
rownames(scorm1) <- samplessubS[orderSeq]
scaleyellowred <- colorRampPalette(c("red", "lightyellow"), space = "rgb")(100)
heatmap.2(scorm1[orderSeq,], dendrogram="none", 
          Rowv=FALSE, Colv=FALSE,
          col = scaleyellowred,
          labRow = "",
          RowSideColors = colcolssubS[orderSeq], 
          margins = c(10, 3),  trace = "none",   
          key=T, key.xlab="1 - Spearman", key.title ="",
          density.info = "none")
```

save figure

```{r, eval=FALSE}
pdf("/Users/Zhixiang/Dropbox/CEGS/figures_Zhixiang/heatmap_hclust_landmark.pdf", width=10, height=10)
heatmap.2(scorm1[orderSeq,], dendrogram="none", 
          Rowv=FALSE, Colv=FALSE,
          col = scaleyellowred,
          labRow = "",
          RowSideColors = colcolssubS[orderSeq], 
          margins = c(10, 3),  trace = "none",   
          key=T, key.xlab="1 - Spearman", key.title ="", keysize = 1.2,
          density.info = "none")
dev.off()
```


## landmap and top peak plot for mahdi

```{r}
library(readr)
data <- read_csv("C:/Users/Zhixiang/Dropbox/tmp/Landmark/Landmark_reorder.csv", col_names = FALSE)
data <- as.matrix(data)
sampleinfor <- read_delim("C:/Users/Zhixiang/Dropbox/tmp/Landmark/KnnWithIntial.txt", "\t", escape_double = FALSE, col_names = FALSE, trim_ws = TRUE)
clusters <- sampleinfor[,2][[1]]
samples <- sampleinfor[,3][[1]]
```

```{r}
cells <- c("K562", "GM12878", "HL-60", "BJ", "TF-1", "H1")
cols <- rep(NA, length(samples))
for (i in 1:length(cells)){
  cell <- cells[i]
  cols[which(samples==cell)] <- brewer.pal(6, "Dark2")[i]
}
scaleyellowred <- colorRampPalette(c("lightyellow", "red"), space = "rgb")(100)

dataThres <- data
dataThres[which(dataThres>=6)] <- 6
#tt <- sample(ncol(dataThres),500)
#tt <- tt[order(tt)]
heatmap.2(dataThres, dendrogram="none", col = scaleyellowred, na.color="yellow",
          labRow = "", labCol = "",
          ## row and column order
          Rowv=FALSE, Colv = FALSE,  
          ## row and column side color
          #ColSideColors = colcolssubS
          RowSideColors = cols, 
          margins = c(10, 3),  trace = "none",   
          ## color key
          key=F, key.xlab="1 - Spearman", key.title="", keysize=1.2, 
          density.info = "none")
```

```{r}
col1 = hclust(as.dist(cor(dataThres[,tt], method="spearman")), method="average")
heatmap.2(dataThres[,tt], dendrogram="column", col = scaleyellowred, na.color="yellow",
          labRow = "", labCol = "",
          ## row and column order
          Rowv=FALSE, Colv = col1,  
          ## row and column side color
          #ColSideColors = colcolssubS
          RowSideColors = cols, 
          margins = c(10, 3),  trace = "none",   
          ## color key
          key=F, key.xlab="1 - Spearman", key.title="", keysize=1.2, 
          density.info = "none")
```

save figure

```{r, eval=FALSE}
pdf("/Users/Zhixiang/Dropbox/CEGS/figures_Zhixiang/heatmap_toppeaks.pdf", width=10, height=10)
heatmap.2(dataThres, dendrogram="none", col = scaleyellowred, na.color="yellow",
          labRow = "", labCol = "",
          ## row and column order
          Rowv=FALSE, Colv = FALSE,  
          ## row and column side color
          #ColSideColors = colcolssubS
          RowSideColors = cols, 
          margins = c(10, 3),  trace = "none",   
          ## color key
          key=F, key.xlab="1 - Spearman", key.title="", keysize=1.2, 
          density.info = "none")
dev.off()
```

save figure, black

```{r, eval=FALSE}
scaleyellowblack <- colorRampPalette(c("lightyellow", "black"), space = "rgb")(100)
pdf("/Users/Zhixiang/Dropbox/CEGS/figures_Zhixiang/heatmap_toppeaksBlack.pdf", width=10, height=10)
heatmap.2(dataThres, dendrogram="none", col = scaleyellowblack, na.color="yellow",
          labRow = "", labCol = "",
          ## row and column order
          Rowv=FALSE, Colv = FALSE,  
          ## row and column side color
          #ColSideColors = colcolssubS
          RowSideColors = cols, 
          margins = c(10, 3),  trace = "none",   
          ## color key
          key=F, key.xlab="1 - Spearman", key.title="", keysize=1.2, 
          density.info = "none")
dev.off()
```

Load the data with more cells

```{r, eval=FALSE}
data <- read_csv("C:/Users/Zhixiang/Data/CellLines6/updated/ForeGroundMore.csv", col_names = FALSE)
data <- as.matrix(data)
```

```{r}
samples <- read_delim("C:/Users/Zhixiang/Data/CellLines6/updated/SampleOrderMore.txt", 
    "\t", escape_double = FALSE, col_names = FALSE, trim_ws = TRUE)
samples <- samples[,2][[1]]
table(samples)
```

```{r}
distS <- 1-cor(data, method="spearman")
resultpamS <- pam(x=distS, k=6, diss = TRUE)
clusterS <- resultpamS$clustering
cells <- c("K562", "GM12878", "HL-60", "BJ", "TF-1", "H1")
```

check clusters. 

```{r}
clustertmp <- clusterS
clusterMembertmp <- matrix(0, nrow=6, ncol=6)
for (i in 1:length(unique(clustertmp))){
  tmp <- samples[which(clustertmp==i)]
  for (j in 1:length(cells)){
    cell <- cells[j]
    clusterMembertmp[i, j] <- sum(tmp==cell)
  }
}
## rearrange it
clusterMembertmp <- clusterMembertmp[apply(clusterMembertmp, 2, which.max),]
row.names(clusterMembertmp) <- paste("cluster", 1:6)
colnames(clusterMembertmp) <- cells
clusterMemberS <- clusterMembertmp
print(clusterMemberS)
```

Agglomerative step afterwards

```{r}
ncenters <- length(unique(clusterS))
clustercenter <- c()
  for (i in 1:ncenters){
    tmp <- which(clusterS==i)
    if (length(tmp)==1){
        clustercenter <- cbind(clustercenter, data[,tmp]  )  
    } else {
        clustercenter <- cbind(clustercenter, rowSums(data[,tmp])  ) 
    }
}
scor <- cor(data, clustercenter, method="spearman")
clusterSA <- apply(scor, 1, which.max) 
```

check clusters. 
```{r}
clustertmp <- clusterSA
clusterMembertmp <- matrix(0, nrow=6, ncol=6)
for (i in 1:length(unique(clustertmp))){
  tmp <- samples[which(clustertmp==i)]
  for (j in 1:length(cells)){
    cell <- cells[j]
    clusterMembertmp[i, j] <- sum(tmp==cell)
  }
}
## rearrange it
clusterMembertmp <- clusterMembertmp[apply(clusterMembertmp, 2, which.max),]
row.names(clusterMembertmp) <- paste("cluster", 1:6)
colnames(clusterMembertmp) <- cells
clusterMemberS <- clusterMembertmp
print(clusterMemberS)
```

## use the top peaks for PAM

```{r}
caloverlap <- function(x, y, top){
  xtmp <- 0*x
  ytmp <- 0*y
  xtmp[order(x, decreasing=T)[1:top]] <- 1  
  ytmp[order(y, decreasing=T)[1:top]] <- 1 
  sum((xtmp+ytmp)==2)/top
}

getdataBin <- function(x, top){
  xtmp <- 0*x
  xtmp[order(x, decreasing=T)[1:top]] <- 1  
  return(xtmp)
}

calBinDist <- function(column, data, top){
  return(colSums(data*column)/top)
}
```

the version with 946 cells. using the top peaks does not work well

```{r}
load("C:/Users/Zhixiang/Data/CellLines6/updated/ForeGround.rda")
samples <- read_delim("C:/Users/Zhixiang/Data/CellLines6/updated/SampleOrder.txt", 
    "\t", escape_double = FALSE, col_names = FALSE, trim_ws = TRUE)
samples <- samples[,2][[1]]
table(samples)
```

```{r}
top <- 1000
dataBin <- apply(data, 2, getdataBin, top)
distTop <- 1-crossprod(dataBin)/top
resultpamS <- pam(x=distTop, k=6, diss = TRUE)
clusterS <- resultpamS$clustering
cells <- c("K562", "GM12878", "HL-60", "BJ", "TF-1", "H1")
```

check clusters. 

```{r}
clustertmp <- clusterS
clusterMembertmp <- matrix(0, nrow=6, ncol=6)
for (i in 1:length(unique(clustertmp))){
  tmp <- samples[which(clustertmp==i)]
  for (j in 1:length(cells)){
    cell <- cells[j]
    clusterMembertmp[i, j] <- sum(tmp==cell)
  }
}
## rearrange it
clusterMembertmp <- clusterMembertmp[apply(clusterMembertmp, 2, which.max),]
row.names(clusterMembertmp) <- paste("cluster", 1:6)
colnames(clusterMembertmp) <- cells
clusterMemberS <- clusterMembertmp
print(clusterMemberS)
```

experiment with merging

```{r, eval=FALSE}
data <- read_csv("C:/Users/Zhixiang/Data/CellLines6/updated/ForeGroundMore.csv", col_names = FALSE)
data <- as.matrix(data)
samples <- read_delim("C:/Users/Zhixiang/Data/CellLines6/updated/SampleOrderMore.txt", 
    "\t", escape_double = FALSE, col_names = FALSE, trim_ws = TRUE)
samples <- samples[,2][[1]]
table(samples)
cells <- c("K562", "GM12878", "HL-60", "BJ", "TF-1", "H1")
```

pick a number of cells for each type to merge and then assign cluster membership based on the merged samples

```{r}
set.seed(123)
num <- 50
labnum <- c()
clustercenter <- c()
for (cell in cells){
    tmp <- sample(which(samples==cell), num)
    labnum <- c(labnum, tmp)
    if (length(tmp)==1){
        clustercenter <- cbind(clustercenter, data[,tmp]  )  
    } else {
        clustercenter <- cbind(clustercenter, rowSums(data[,tmp])  ) 
    }
}

scor <- cor(data, clustercenter, method="spearman")
cluster <- apply(scor, 1, which.max)
```

```{r}
clustertmp <- cluster[labnum]
samplestmp <- samples[labnum]
clusterMembertmp <- matrix(0, nrow=6, ncol=6)
for (i in 1:length(unique(clustertmp))){
  tmp <- samplestmp[which(clustertmp==i)]
  for (j in 1:length(cells)){
    cell <- cells[j]
    clusterMembertmp[i, j] <- sum(tmp==cell)
  }
}
## rearrange it
clusterMembertmp <- clusterMembertmp[apply(clusterMembertmp, 2, which.max),]
row.names(clusterMembertmp) <- paste("cluster", 1:6)
colnames(clusterMembertmp) <- cells
clusterMemberS <- clusterMembertmp
print(clusterMemberS)
```

```{r}
clustertmp <- cluster[-labnum]
samplestmp <- samples[-labnum]
clusterMembertmp <- matrix(0, nrow=6, ncol=6)
for (i in 1:length(unique(clustertmp))){
  tmp <- samplestmp[which(clustertmp==i)]
  for (j in 1:length(cells)){
    cell <- cells[j]
    clusterMembertmp[i, j] <- sum(tmp==cell)
  }
}
## rearrange it
clusterMembertmp <- clusterMembertmp[apply(clusterMembertmp, 2, which.max),]
row.names(clusterMembertmp) <- paste("cluster", 1:6)
colnames(clusterMembertmp) <- cells
clusterMemberS <- clusterMembertmp
print(clusterMemberS)
```

Try it for multiple runs


```{r}
set.seed(123)
nrun <- 20
nums <- c(3, 5, 10, 20, 50)
accNumA <- matrix(nrow=length(nums), ncol=nrun)
accOtherA <- matrix(nrow=length(nums), ncol=nrun)
accAllA <- matrix(nrow=length(nums), ncol=nrun)
for (k in 1:length(nums)){
  num <- nums[k]
  accNum <- c()
  accOther <- c()
  accAll <- c()
  for (run in 1:nrun){
    print(run)
    labnum <- c()
    clustercenter <- c()
    for (cell in cells){
        tmp <- sample(which(samples==cell), num)
        labnum <- c(labnum, tmp)
        if (length(tmp)==1){
            clustercenter <- cbind(clustercenter, data[,tmp]  )  
        } else {
            clustercenter <- cbind(clustercenter, rowSums(data[,tmp])  ) 
        }
    }
    
    scor <- cor(data, clustercenter, method="spearman")
    cluster <- apply(scor, 1, which.max)
    
    clustertmp <- cluster[labnum]
    samplestmp <- samples[labnum]
    clusterMembertmp <- matrix(0, nrow=6, ncol=6)
    for (i in 1:length(unique(clustertmp))){
      tmp <- samplestmp[which(clustertmp==i)]
      for (j in 1:length(cells)){
        cell <- cells[j]
        clusterMembertmp[i, j] <- sum(tmp==cell)
      }
    }
    ## rearrange it
    clusterMembertmp <- clusterMembertmp[apply(clusterMembertmp, 2, which.max),]
    row.names(clusterMembertmp) <- paste("cluster", 1:6)
    colnames(clusterMembertmp) <- cells
    clusterMemberS <- clusterMembertmp
    cN <- sum(diag(clusterMemberS))
    accNum <- c(accNum, cN/sum(clusterMemberS))
    
    
    clustertmp <- cluster[-labnum]
    samplestmp <- samples[-labnum]
    clusterMembertmp <- matrix(0, nrow=6, ncol=6)
    for (i in 1:length(unique(clustertmp))){
      tmp <- samplestmp[which(clustertmp==i)]
      for (j in 1:length(cells)){
        cell <- cells[j]
        clusterMembertmp[i, j] <- sum(tmp==cell)
      }
    }
    ## rearrange it
    clusterMembertmp <- clusterMembertmp[apply(clusterMembertmp, 2, which.max),]
    row.names(clusterMembertmp) <- paste("cluster", 1:6)
    colnames(clusterMembertmp) <- cells
    clusterMemberS <- clusterMembertmp
    cO <- sum(diag(clusterMemberS))
    accOther <- c(accOther, cO/sum(clusterMemberS))
    
    accAll <- c(accAll, (cN+cO)/ncol(data))
  }
  accNumA[k,] <- accNum
  accOtherA[k,] <- accOther
  accAllA[k,] <- accAll
}
```

boxplot

```{r}
boxplot(t(rbind(accNumA, accOtherA, accAllA)))
```

## try weighted K-medoid


