---
title: "scATAC_GM_K562"
author: "Zhixiang Lin"
date: "2/2/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GM, 5000 regions

```{r echo=TRUE, message=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
library(rafalib)
GM <- as.matrix(read.table("~/Documents/collaboration/mahdi/scATAC/GM_5000.txt"))

check_bimodal <- function(x, perc1, perc2, thres1, thres2){
  (((sum(x<=thres1)/length(x)) >= perc1) & ((sum(x>=thres2)/length(x)) >= perc2)) + 0
}


labs_GM <- apply(GM, 1, check_bimodal, 0.3, 0.3, 1, 100)
row.names(GM) <- 1:5000
mypar(mfrow=c(1,1))
heatmap(t(GM[which(labs_GM==1),]))
```

```{r echo=TRUE, message=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
mypar(mfrow=c(2,2))
hist(GM[7,], breaks=30)
hist(GM[393,], breaks=30)
hist(GM[4,], breaks=30)
hist(GM[68,], breaks=30)
```

## GM, 5000 regions, check robustness of the clusters

Binarize the data matrix, fold-change>50 open(1); otherwise close(0)

```{r echo=TRUE, message=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
data <- t(GM[which(labs_GM==1),])
data_thres <- data
thres <- 50
data_thres[which(data>=thres)] <- 1
data_thres[which(data<thres)] <- 0

kmean <- kmeans(data, centers=2, nstart=500)$cluster
kmean_thres <- kmeans(data_thres, centers=2, nstart=500)$cluster
# check cluster consistency
table(kmean)
table(kmean_thres)
table(kmean_thres[which(kmean==1)])
table(kmean_thres[which(kmean==2)])
purity <- (max(table(kmean_thres[which(kmean==1)])) + max(table(kmean_thres[which(kmean==2)])))/nrow(data)
print(paste("purity: ", purity))
# random cluster
kmean_thres_random <- sample(kmean_thres)
table(kmean_thres_random[which(kmean==1)])
table(kmean_thres_random[which(kmean==2)])
purity_random <- (max(table(kmean_thres_random[which(kmean==1)])) + max(table(kmean_thres_random[which(kmean==2)])))/nrow(data)
print(paste("purity random: ", purity_random))
```

Heatmap after binarization

```{r echo=TRUE, message=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
mypar(mfrow=c(1,1))
heatmap(data_thres, scale = "none", col = c("white", "black"))
```

## GM, 100,000 regions

Check overall distribution of fold change

```{r echo=TRUE, message=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
dataA <- read.csv("~/Documents/collaboration/mahdi/scATAC/GM12878.csv", header = FALSE)
dataA <- as.matrix(dataA)
boxplot(log2(dataA[,sample(ncol(dataA), 10)]+1), ylab="log fold change")
```

What does the distribution for the non-zero log fold change look like per sample?

```{r echo=TRUE, message=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
set.seed(123)
mypar(mfrow=c(3,3))
nsam <- 9
for (sam in 1:nsam){
  lab <- sample(ncol(dataA), 1)
  tmp <- log2(dataA[,lab]+1)
  hist(tmp[tmp!=0], breaks=200, xlab=paste("sample", lab), main="")
}
```

Number of non-zero fold change per sample

```{r echo=TRUE, message=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
fc <- apply(dataA, 2, function(x){sum(x!=0)})
summary(fc)
mypar(mfrow=c(1, 1))
boxplot(fc, ylab="distribution for the number of fold change (exclude 0s), per sample")
```

Binarize the data matrix, boxplot for the number of peaks per sample

```{r echo=TRUE, message=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
thres <- 50
dataA_bin <- dataA
dataA_bin[dataA>=thres] <- 1
dataA_bin[dataA<thres] <- 0
peak_numberA <- colSums(dataA_bin)
### two problematic samples, sample 301 and 316
which(is.na(peak_numberA))
boxplot(peak_numberA)
```

Binarize the data matrix, boxplot for the number of peaks per sample, stratified by the cluster label

```{r echo=TRUE, message=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
mypar(mfrow=c(1,2))
boxplot(peak_numberA~kmean, xlab="cluster label, top bimodal regions")
boxplot(peak_numberA~kmean_thres, xlab="cluster label, top bimodal regions, binarize")
```

We see that in some samples, more regions have very large fold changes. We need to some normalization/adjustment. For a sample, if the sequencing depth is shallow, it is less likely that there will be peaks with large fold change?

For each sample, pick the top regions. Then do cluster based on the top regions

```{r echo=TRUE, message=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
top <- 1000
# samples with less than top regions 
dataA_sub <- dataA[,which(fc>top)]
# the labels
labs <- apply(dataA_sub, 2, function(x, top){(order(x, decreasing = T))[1:top]}, top)
# binarize by picking the top regions
dataA_sub_bin <- matrix(0, nrow=nrow(dataA_sub), ncol=ncol(dataA_sub))
for (i in 1:ncol(dataA_sub)){
  dataA_sub_bin[labs[,i],i] <- 1
}
dataA_sub_binRS <- rowSums(dataA_sub_bin)
sum(dataA_sub_binRS==0)
dataA_sub_bin_random <- apply(dataA_sub_bin, 2, sample)
dataA_sub_bin_randomRS <- rowSums(dataA_sub_bin_random)
sum(dataA_sub_bin_randomRS==0)
mypar(mfrow=c(1,2))
hist(dataA_sub_binRS, breaks=200)
hist(dataA_sub_bin_randomRS, breaks=200)
```

```{r echo=TRUE, message=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
# open in more than perc of the samples
perc <- 0.15
sum(dataA_sub_binRS>=perc*ncol(dataA_sub_bin))
heatmap(t(dataA_sub_bin[which(dataA_sub_binRS>=perc*ncol(dataA_sub_bin)),]), scale = "none", col = c("white", "black"))
```

## GM, exclude low background reads

Define distance based on the percentage of overlap regions, where both samples have reliable background estimate. This part runs on cluster.

```{r, eval=FALSE, echo=TRUE, message=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
dataAll <- read.csv("/scratch/PI/whwong/mzamani/Shining/GM12878.csv", header = FALSE)
dataAll <- as.matrix(dataAll)

cutoffs <- seq(20, 100, 20)
thress <- c(0, 10, 20, 50)

for (thres in thress){
  for (cutoff in cutoffs){
    fn <- paste("cutoff_", cutoff, "_thres_", thres, ".rda", sep="")
    fn_count <- paste("cutoff_", cutoff, "_thres_", thres, "_numregions.rda", sep="")
    datasub <- read.csv(paste("/scratch/PI/whwong/mzamani/Shining/GM", cutoff, ".csv", sep=""),      header = FALSE)
    datasub <- as.matrix(datasub)

    bad_background <- (datasub!=dataAll) + 0
    distm <- matrix(0, nrow=ncol(dataAll), ncol=ncol(dataAll))
    numregions <- matrix(0, nrow=ncol(dataAll), ncol=ncol(dataAll))
    for (i in 1:(ncol(dataAll)-1)){
      print(i)
      for (j in (i+1):ncol(dataAll)){
        filter1 <- datasub[,i]<=thres & datasub[,j]<=thres
        filter2 <- bad_background[,i]==1 | bad_background[,j]==1
        d1 <- datasub[-which(filter1 | filter2),i]
        d2 <- datasub[-which(filter1 | filter2),j]
        distm[i,j] <- distm[j,i] <- 1-(sum(d1>0 & d2>0)/length(d1))
        numregions[i,j] <- numregions[j,i] <- length(d1)
      }
    }
    save(distm, file=fn)
    save(numregions, file=fn_count)
  }
}
```

Plot the results from the cluster, 

```{r, eval=FALSE, echo=TRUE, message=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
cutoffs <- seq(20, 100, 20)
thress <- c(0, 10, 20, 50)
for (thres in thress){
  for (cutoff in cutoffs){
    fn <- paste("~/Documents/collaboration/mahdi/scATAC/cutoff_", cutoff, "_thres_", thres, ".rda", sep="")
    load(fn)
    # exclude the rows/columns with NA
    if (sum(is.na(rowSums(distm)))){
      distm <- distm[which(!is.na(rowSums(distm))),  which(!is.na(rowSums(distm)))]  
    }
    pdf(paste("~/Documents/collaboration/mahdi/scATAC/cutoff_", cutoff, "_thres_", thres, "_hclust.pdf", sep=""))
    plot(hclust(as.dist(distm)))
    dev.off()
    
    pdf(paste("~/Documents/collaboration/mahdi/scATAC/cutoff_", cutoff, "_thres_", thres, "_mds.pdf", sep=""))
    plot(cmdscale(distm, k = 2))
    dev.off()
  }
}

## power of distm
library(gplots)
cutoffs <- seq(20, 100, 20)
thress <- c(0, 10, 20, 50)
for (thres in thress){
  for (cutoff in cutoffs){
    fn <- paste("~/Documents/collaboration/mahdi/scATAC/cutoff_", cutoff, "_thres_", thres, ".rda", sep="")
    load(fn)
    # exclude the rows/columns with NA
    if (sum(is.na(rowSums(distm)))){
      labs <- (1:ncol(distm))[which(!is.na(rowSums(distm)))]
      distm <- distm[which(!is.na(rowSums(distm))),  which(!is.na(rowSums(distm)))]  
    }
    colnames(distm) <- labs
    rownames(distm) <- labs
    power <- 3
    pdf(paste("~/Documents/collaboration/mahdi/scATAC/cutoff_", cutoff, "_thres_", thres, "_power_", power, "_heatmap.pdf", sep=""))
    heatmap.2(distm^power, Colv="Rowv",trace='none', main=paste("nsam=", ncol(distm), ", power=", power, sep=""))
    dev.off()
    power <- 1
        pdf(paste("~/Documents/collaboration/mahdi/scATAC/cutoff_", cutoff, "_thres_", thres, "_power_", power, "_heatmap.pdf", sep=""))
    heatmap.2(distm^power, Colv="Rowv",trace='none', main=paste("nsam=", ncol(distm), ", power=", power, sep=""))
    dev.off()
  }
}

## check the number of regions for each pair of samples
cutoffs <- seq(20, 100, 20)
thress <- c(0, 10, 20, 50)
for (thres in thress){
  for (cutoff in cutoffs){
    fn <- paste("~/Documents/collaboration/mahdi/scATAC/cutoff_", cutoff, "_thres_", thres, ".rda", sep="")
    load(fn)
    fnnum <- paste("~/Documents/collaboration/mahdi/scATAC/cutoff_", cutoff, "_thres_", thres, "_numregions.rda", sep="")
    load(fnnum)
    
    distmUP <- distm[which(upper.tri(distm))]
    numregionsUP <- numregions[which(upper.tri(numregions))]
    # exclude the rows/columns with NA
    if (sum(is.na(rowSums(distm)))){
      labs <- (1:ncol(distm))[which(!is.na(rowSums(distm)))]
      distm <- distm[labs, labs]  
      numregions <- numregions[labs, labs] 
    }
    colnames(distm) <- labs
    rownames(distm) <- labs
    power <- 3
    
    
    
    a <- heatmap.2(distm^power, Colv="Rowv",trace='none', main=paste("nsam=", ncol(distm), ", power=", power, sep=""))
    
    cluster <- cutree(as.hclust(a$rowDendrogram), k = 2)
    
    num22 <- numregions[which(cluster==2),which(cluster==2)]
    num11 <- numregions[which(cluster==1),which(cluster==1)]
    
    hist(log10(num22[which(upper.tri(num22))]+1), breaks=200, freq=F, xlim=c(0, 5))
    hist(log10(num11[which(upper.tri(num11))]+1), breaks=200, freq=F, add=T, col="red")
    

  }
}
```

```{r echo=TRUE, message=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
col.names(distm) <- row.names(distm) <- 1:ncol(distm)
hclust(dist)
```

```{r echo=TRUE, message=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
hist(numregions[which(upper.tri(numregions))], breaks=100)
```

## K562, 5000 regions
```{r echo=TRUE, message=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
K562 <- as.matrix(read.table("~/Documents/collaboration/mahdi/scATAC/K562_5000.txt"))
labs_K562 <- apply(K562, 1, check_bimodal, 0.3, 0.1, 1, 20)
sum(labs_K562)
row.names(K562) <- 1:5000
mypar(mfrow=c(1,1))
heatmap(t(K562[which(labs_K562==1),]))
```

```{r echo=TRUE, message=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
mypar(mfrow=c(2,2))
hist(K562[989,], breaks=30)
hist(K562[1163,], breaks=30)
hist(K562[3531,], breaks=30)
hist(K562[2754,], breaks=30)
```

## K562, 5000 regions, check robustness of the clusters

Binarize the data matrix, fold-change>20 open(1); otherwise close(0)

```{r echo=TRUE, message=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
data <- t(K562[which(labs_K562==1),])
data_thres <- data
thres <- 20
data_thres[which(data>=thres)] <- 1
data_thres[which(data<thres)] <- 0

kmean <- kmeans(data, centers=2, nstart=500)$cluster
kmean_thres <- kmeans(data_thres, centers=2, nstart=500)$cluster
# check cluster consistency
table(kmean)
table(kmean_thres)
table(kmean_thres[which(kmean==1)])
table(kmean_thres[which(kmean==2)])
purity <- (max(table(kmean_thres[which(kmean==1)])) + max(table(kmean_thres[which(kmean==2)])))/nrow(data)
print(paste("purity: ", purity))
# random cluster
kmean_thres_random <- sample(kmean_thres)
table(kmean_thres_random[which(kmean==1)])
table(kmean_thres_random[which(kmean==2)])
purity_random <- (max(table(kmean_thres_random[which(kmean==1)])) + max(table(kmean_thres_random[which(kmean==2)])))/nrow(data)
print(paste("purity random: ", purity_random))
```

Heatmap after binarization

```{r echo=TRUE, message=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
mypar(mfrow=c(1,1))
heatmap(data_thres, scale = "none", col = c("white", "black"))
```

